---
title: "Making an .sps file"
author: "Jacob Kaplan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Making an .sps file}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
Most times you will deal with a .dat+.sps file pair, you will be given
the .sps file already made. However, in rare cases - such as working with
FBI data - you need to make you own based on a PDF guide provided. 

The function `make_sps_setup` makes that a bit easier by doing all the 
formatting needed, meaning you just need to provide the info for each
section. It will create a .sps setup file and save it in your working directory.

There are two parameters absolutely needed - `file_name` and `col_positions` - as 
well as four optional parameters. 
  
* `file_name` - Name of the file to be saved (e.g. "setup_file1"). There is no need to put the .sps extention in the file name.
* `col_positions` - Either a vector of strings indicating the start and end position of each column (e.g. "1-3", "4-5") or a vector of the widths of the columns (e.g. 3, 2).
* `col_names` - A vector of names for the columns. If none are provided, will automatically  create names based on column number (e.g. V1, V2, V3).
* `col_labels` - A vector of labels for the columns. These are often longer and more descriptive than the col_names. These are the values used as column names if real_names = TRUE in reading in the data.
* `value_labels` - A named vector with the value first, then an ' = ' then the label. Each
  new column should have the column named followed by ' ='.
* `missing_values` - A vector of strings with the column name followed by the values to be
  replaced by NA.
  
  
## file_name
`file_name` is simply the string of the name you want the file to be saved as, such as "setup_file_example". Including the ".sps" to the end is not required.
  
## col_positions
  
```{r}
col_positions <- c("1-1", "2-3", "4-5", "6-11")
```

## col_names

```{r}
col_names <- c("var1", "var2", "var3", "var4")
```

## col_labels

```{r}
col_labels <- c("version_number", "victim_sex", "victim_race", "state")
```

## value_labels

```{r}
value_labels <- c("victim_sex = ",
                  "MA = male",
                  "FE = female",
                  "UN = unknown",
                  "victim_race = ",
                  "WH = white",
                  "BL = black",
                  "IA = american indian or alaskan native",
                  "AS = asian",
                  "UN = unknown")
```

## missing_values

```{r}
missing_values <- c("victim_sex", 
                    "-9", 
                    "-8", 
                    "victim_race",
                    "-8")
```

## Putting it all together

```{r eval = FALSE}
asciiSetupReader::make_sps_setup(file_name = "setup_file_example",
                                 col_positions  = col_positions,
                                 col_names      = col_names,
                                 col_labels     = col_labels,
                                 value_labels   = value_labels,
                                 missing_values = missing_values)
```

