---
title: "Parsing an .sps or .sas file"
author: "Jacob Kaplan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Making an .sps file}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


The `parse_setup` function is what this packages uses to convert the .sps or .sas setup files into an usable format for R. 

```{r}
# Using the example .sps setup file included with the package
sps_name <- system.file("extdata", "example_setup.sps",
                          package = "asciiSetupReader")
example <- parse_setup(sps_name)
```

## setup

The first object in the returned list is a data.frame with 4 columns and as many rows as there are columns in the data. 

```{r}
head(example$setup)
```


## value_labels

```{r}
example$value_labels[1:3]
```
There is one named vector for each column in the data that has value labels. We can see how many there are using `length()`.

```{r}
length(example$value_labels)
```

## missing

The final object in the list if a 

```{r}
head(example$missing)
```

